---
title: "Indicadores de impacto y resultados mediante la CASEN"
output: unhcrdown::pptx_slides
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.cap = TRUE, message = FALSE, warning = FALSE)
library(officedown)
library(unhcrthemes)
library(fontawesome)
library(haven)
library(dplyr)
library(scales)
library(tidyverse)

# turn off the automatic use of showtext functionality, so that the dml function can works properly
showtext::showtext_auto(FALSE) 

casen2022 <- read_sav("data-raw/Casen_2022.sav")
```


```{r setup2, include=FALSE}
# names(casen2022 )

## Explore all content
#  casen2022  |>  sjPlot::view_df()

casen2022 <- casen2022 |>
  dplyr::mutate(pop = dplyr::case_when(
    r1b == 3 & 
      r1b_pais_esp_cod == 513 #& 
      ## r1cp  arrived aftr 2018/....
      #r1cp %in% c(1,2) 
      ~ "Venezuelans",
    r1b== 1 | r1b == 2  ~ "Chileans",
    TRUE ~ "Others"))

 

```

----

# Encuesta de Caracterización Socioeconómica Nacional

La [Encuesta Casen](https://observatorio.ministeriodesarrollosocial.gob.cl/encuesta-casen-2022) es una encuesta a hogares, de carácter transversal y multipropósito, realizada por el Ministerio de Desarrollo Social y Familia (antes Ministerio de Planificación y Cooperación). Ha sido levantada de manera regular en el país desde 19872. Hasta la fecha, se han realizado 15 versiones de la Encuesta en los años 1987, 1990, 1992, 1994, 1996, 1998, 2000, 2003, 2006, 2009, 2011, 2013, 2015, 2017, 2020 y [2022](https://observatorio.ministeriodesarrollosocial.gob.cl/storage/docs/casen/2022/Base%20de%20datos%20Casen%202022%20SPSS.sav.zip).

----

# Objetivos Encuesta CASEN:

- Conocer la situación de pobreza por ingresos de las personas y los hogares, así como la distribución del ingreso de los hogares.
- Identificar carencias de la población en las áreas de educación, salud, vivienda, trabajo e ingresos.
- Evaluar brechas de pobreza por ingresos y carencias entre distintos grupos de la población como niños, niñas y adolescentes; jóvenes; personas mayores; mujeres; pueblos indígenas; migrantes; entre otros.
- Evaluar brechas de pobreza por ingresos y carencias entre zonas urbanas y rurales, y entre territorios.
- Estimar cobertura, focalización y distribución del gasto fiscal de los principales subsidios monetarios de alcance nacional entre los hogares,  para evaluar su impacto.

----

# Impact Area 2

```{r layout='Title and Content', ph=officer::ph_location_type(type="body")}

# 2.1 Proportion of PoC living below the national poverty line.
#----2.3 Proportion of PoC with access to health services----
rbm_indicators <- 
  c(rbm2.1 = "2.1 Proportion of PoC living below the national poverty line",
    rbm2.3 = "2.3 Proportion of PoC with access to health services")

#impact area 2
indicators1 <-   casen2022 %>% 
  # transmute function returns the same new variable as mutate.
  # However, it does not retain our original data!
  transmute(pop = pop,
            # 2.1 Proportion of PoC living below the national poverty line
            rbm2.1 = pobreza== 1 | pobreza== 2,
            # 2.3 Proportion of PoC with access to health services
            rbm2.3 = if_else(s16 == 1, 
                             is.na (s18), 
                             NA),
           region = region, 
           wt = expr)

  ## get all indicators joined
ind <-  indicators1  %>%
  ## Remove the neither colombian nor venezuelan
  #filter(pop != "Others") %>% 
  ## transform in a weighted survey object
  srvyr::as_survey_design(strata = region, 
                          weights = wt) %>% 
  ## Compile indicators by group
  dplyr::group_by(pop) %>% 
  dplyr::summarize(
    # Apply functions across multiple columns
    dplyr::across(
      ## select all variable name based on 
      ## name pattern
      dplyr::contains("rbm"), 
      srvyr::survey_mean, 
      # Report variability as confidence interval ("ci")
      vartype = "ci", 
      # missing values should be dropped
      na.rm = TRUE,
      ## append a suffix to all column name to 
      # distinguish the mai estimation from CI
      .names = "{.col}_est"))

ind_chart <- ind   %>%
  ## pivot from wide to long using variable pop
  pivot_longer(-pop, 
               names_to = c("ind", ".value"), 
               names_pattern = "(.+?)_(.+)")


plot <- ggplot( data = ind_chart,
                aes( x= est, 
                     y = fct_rev(ind))) +
  geom_pointrange(position = position_dodge(width = 0.2), aes(xmin = est_low, 
                      xmax = est_upp, 
                      color = pop), 
                  size = .75) +
  ggrepel::geom_text_repel(
    aes(label = scales::label_percent(.1)(est)),
    box.padding = .5,
    min.segment.length = 1,
    point.padding = .8  )  +
  scale_x_continuous(labels = scales::label_percent(), 
                     limits = c(0, 1), 
                     breaks = c(0, .5, 1)) +
  scale_color_manual(values = c(Chileans = "black", Venezuelans = "#0072BC", Others = "grey50"))  +
## Add labels for each calculated indicators 
  scale_y_discrete(labels = str_wrap(rev(rbm_indicators),40))   + 
  labs(x = "Estimate", y = "",
       title = "Impact Area 2: Realizing Rights in Safe Environments",
       caption = glue::glue("Encuesta de Caracterización Socioeconómica Nacional (CASEN) 2022"
                            )) +
  unhcrthemes::theme_unhcr(font_size = 15) +
  theme(panel.grid.minor.x = element_blank())

 
# turn off the automatic use of showtext functionality, so that the dml function can works properly
showtext::showtext_auto(FALSE) 
rvg::dml(ggobj = plot, fonts = list(serif = 'Lato'))

```

----

# Impact Area 3


```{r layout='Title and Content', ph=officer::ph_location_type(type="body")}

rbm_indicators <-   c(rbm3.2a = "3.2a Proportion of PoC enrolled in primary education",
    rbm3.2b = "3.2b Proportion of PoC enrolled in secundary education")

#impact area 2
indicators1 <-   casen2022 %>% 
  # transmute function returns the same new variable as mutate.
  # However, it does not retain our original data!
  transmute(pop = pop,
            # 3.2a Proportion of PoC enrolled in primary education
            rbm3.2a = if_else(edad %in% 6:13,
                              #e2==1 & 
                              e6a_asiste %in% c(5,7),
                              NA), 
            # 3.2a Proportion of PoC enrolled in secondary education
            rbm3.2b = if_else (edad %in% 14:18,
                               #e2==1 & 
                               e6a_asiste %in% c(9,11),
                               NA), 
           region = region, 
           wt = expr)

  ## get all indicators joined
ind <-  indicators1  %>%
  ## Remove the neither colombian nor venezuelan
  #filter(pop != "Others") %>% 
  ## transform in a weighted survey object
  srvyr::as_survey_design(strata = region, 
                          weights = wt) %>% 
  ## Compile indicators by group
  dplyr::group_by(pop) %>% 
  dplyr::summarize(
    # Apply functions across multiple columns
    dplyr::across(
      ## select all variable name based on 
      ## name pattern
      dplyr::contains("rbm"), 
      srvyr::survey_mean, 
      # Report variability as confidence interval ("ci")
      vartype = "ci", 
      # missing values should be dropped
      na.rm = TRUE,
      ## append a suffix to all column name to 
      # distinguish the mai estimation from CI
      .names = "{.col}_est"))

ind_chart <- ind   %>%
  ## pivot from wide to long using variable pop
  pivot_longer(-pop, 
               names_to = c("ind", ".value"), 
               names_pattern = "(.+?)_(.+)")


plot <- ggplot( data = ind_chart,
                aes( x= est, 
                     y = fct_rev(ind))) +
  geom_pointrange(position = position_dodge(width = 0.2), aes(xmin = est_low, 
                      xmax = est_upp, 
                      color = pop), 
                  size = .75) +
  ggrepel::geom_text_repel(
    aes(label = scales::label_percent(.1)(est)),
    box.padding = .5,
    min.segment.length = 1,
    point.padding = .8  )  +
  scale_x_continuous(labels = scales::label_percent(), 
                     limits = c(0, 1), 
                     breaks = c(0, .5, 1)) +
  scale_color_manual(values = c(Chileans = "black", Venezuelans = "#0072BC", Others = "grey50"))  +
## Add labels for each calculated indicators 
  scale_y_discrete(labels = str_wrap(rev(rbm_indicators),40))   + 
  labs(x = "Estimate", y = "",
       title = "Impact Area 3: Empowering Communities and Achieving Gender Equality",
       caption = glue::glue("Encuesta de Caracterización Socioeconómica Nacional (CASEN) 2022"
                            )) +
  unhcrthemes::theme_unhcr(font_size = 15) +
  theme(panel.grid.minor.x = element_blank())

 
# turn off the automatic use of showtext functionality, so that the dml function can works properly
showtext::showtext_auto(FALSE) 
rvg::dml(ggobj = plot, fonts = list(serif = 'Lato'))
 

```
----



# Thank you!


---
